---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# monkeytools

<!-- badges: start -->
[![R-CMD-check](https://github.com/emilio-berti/monkeytools/workflows/R-CMD-check/badge.svg)](https://github.com/emilio-berti/monkeytools/actions)
<!-- badges: end -->

The goal of monkeytools is to provide some utilities functions that I use routinely. Every contribution is welcomed.

## Installation

You can install the development version of monkeytools from [GitHub](https://github.com/emilio-berti/monkeytools) with:

``` r
devtools::install_github("emilio-berti/monkeytools")
```

monkeytools is not released on CRAN.

## Example

### Simple assessments of species distribution models.

Let's first generate some data for a virtual species that never goes above 130 meters.

```{r example}
library(monkeytools)
library(raster)

data(volcano)
r <- raster(volcano)
xy <- xyFromCell(r, sample(1:ncell(r), 100))
x <- extract(r ,xy) #elevation
p <- ifelse(x < 130, 1, 0) #real presence/absence - doesn't like heights
occ <- data.frame(x = xy[, 1], y = xy[, 2], p = p) #combined in a dataframe
head(occ)

# sdm potential output
sdm <- 1 / r + rnorm(ncell(r), 0, .001) #sdm predictions
sdm <- sdm - sdm@data@min
sdm <- sdm / sdm@data@max
```

A very useful summary table is the *confusion matrix*. The confusion table calculates the fraction of correctly predicted occurrences based on observed data. For this, SDM output needs to be binarized.

```{r confusion}
sdm_bin <- sdm
sdm_bin[sdm < .5] <- 0
sdm_bin[sdm >= .5] <- 1
confusion(sdm_bin, occ)
```

The confusion matrix is used in the famous **AUROC** curve, which shows how much some statistics extracted from the confusion matrix change for different binarization thresholds. The area under the AUROC curve (**AUC**) measures how well the SDM fits data better than random (dashed line). There are other (and better) summary statistics for goodness-of-fit of SDMs, but AUC remains one the most adopted.

```{r auroc}
null <- AUROC(sdm, occ, th = seq(0, 1, length.out = 50))
```

And for a very bad SDM:

```{r bad auroc}
values(sdm) <- rnorm(ncell(sdm))
null <- AUROC(sdm, occ, th = seq(0, 1, length.out = 50))
```



<!-- You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>. -->

<!-- You can also embed plots, for example: -->

<!-- ```{r pressure, echo = FALSE} -->
<!-- plot(pressure) -->
<!-- ``` -->

<!-- In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN. -->
